---
title: "Set handling with MAgPIEsets"
author: "Kristine Karstens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with magpiesets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The MAgPIEsets package contains helper functions for reporting and validation of the [MAgPIE](https://github.com/magpiemodel/magpie) model, especially for the preprocessing library *moinput* (see `vignette("moinput")`) and the postprocessing library *magpie4* (on [github](https://github.com/pik-piam/magpie4)).
Since sets evolve over time, the sets are upwards compatible with previous versions.

## Objective
This help will provide some general information about the set management and versioning. It will contain as well help to extend sets or create new items.

### The magpiesets table
The magpiesets package is centered around a *magpiesets table* coming with the package. You can view the table via:

```{r table-sets, eval = FALSE}
View(read.csv(system.file("extdata", "magpiesets.csv", package = "magpiesets")))
```

The table contains in its rows items about 

* commodities, 
* time steps, 
* processing techniques,
* animal waste management systems
* greenhouse gases,
* landuse types,
* sets of commodities and
* product attributes.

These items are allocated to a specific sets by containing the set name within the column of the set.

#### Example:

|                            |human             |female            |male              |
|:---------------------------|:-----------------|:-----------------|:-----------------|
|Alice                       |human             |female            |                  |
|Bob                         |human             |                  |male              |

Since items and set names are often sparse, more detailed names can be found within the *reportingnames table* also coming with the package:

```{r table-names, eval = FALSE}
View(read.csv(system.file("extdata", "reportingnames.csv", package = "magpiesets")))
```

#### Calling a set

Use the `findset` function (help `?findset`) to call set items of a MAgPIE set:
```{r findset1, eval = FALSE}
findset("kall")
```

### Using set versions

Since sets evolve over time, the sets have to be upwards compatible with previous versions.
Therefore the *magpiesets table* contains at the last columns (starting with `default_sep18`) so called *version sets*.

At the moment the following *version sets* are available:

|||
|:---------------|:------------------------------------------------------------------|
| default_sep18:| default MAgPIE 4.0 set with 49 products in 'kall'                  |
| biogas_dec18: | biogas extension including biogas and manure in 'kall'             |
| biochar_jan19 | biochar extension including biochar and pyrolyzing                 |

Please have a look into the *magpiesets table* for more detailed information about item affiliations.

#### Calling a set of a specific version

The `findset` function (help `?findset`) contains the parameter `version`, where you can specify the wanted version:

```{r findset2, eval = FALSE}
findset("kall", version="biogas_dec18")
```

### Expanding the set space

It is possible to extend the *magpiesets table* with both new items and new sets. 

Please note:

* While you have to introduce a new *version set* for new items (new rows in the *magpiesets table*), is it not necessary to do that for new sets (new columns in the *magpiesets table*). 
* Both items and sets have to be added to the *reportingnames table* to provide a meaningful translation.

#### Examples

**Introducing a new item (here *Carol*)**

|                      |human             |female            |male              |default_jul69     |*thirdgender_jan19* |
|:---------------------|:-----------------|:-----------------|:-----------------|:-----------------|:-------------------|
|Alice                 |human             |female            |                  |default_jul69     |*thirdgender_jan19* |
|Bob                   |human             |                  |male              |default_jul69     |*thirdgender_jan19* |
|*Carol*               |*human*           |                  |                  |                  |*thirdgender_jan19* |

Here the new item is added as a row as well as all its affiliations to know sets. Also a new *version set* has to be introduced.

**Introducing a new set (here *diverse*)**

|                         |human       |female      |male        |*diverse*   |default_jul69     |thirdgender_jan19 |
|:------------------------|:-----------|:-----------|:-----------|:-----------|:-----------------|:-----------------|
|Alice                    |human       |female      |            |            |default_jul69     |thirdgender_jan19 |
|Bob                      |human       |            |male        |            |default_jul69     |thirdgender_jan19 |
|Carol                    |human       |            |            |*diverse*   |                  |thirdgender_jan19 |

Here the set just has to be added as a column before the *version sets* at the end of the table.

### More information

The magpiesets package also contains more functionallity not (or minor) touch by the set versioning.
See also:

* `?reportingnames`
* `?reportingReverse`
* `?reporthelper`
* `?summationhelper` 
